---
title: "project DPA"
author: "Priyanka"
date: "2023-04-22"
output:
  word_document: default
  html_document: default
---

```{r}
#Install the pacman package
if(!require(pacman)) install.packages("pacman", repos = "http://cran.us.r-project.org")

#Load the required libraries
#If a package below is missing, p_load will automatically download it from CRAN
pacman::p_load(tidyverse, ggplot2, ggthemes, data.table, lubridate, caret, 
               knitr, scales, treemapify)
pacman::p_load(stringr)
pacman::p_load(dplyr)
```
Data Preparation

```{r}
#Download File
dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)
```

```{r}
#Construct a data frame called 'ratings' by utilizing the 'fread' function from the data.table library
ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\:\\:", 3)

colnames(movies) <- c("movieId", "title", "genres")
```

```{r}
movies <- as.data.frame(movies) %>% 
  mutate(movieId = as.numeric(unique(movieId)),
         title = as.character(title),
         genres = as.character(genres))
movielens <- left_join(ratings, movies, by = "movieId")
```

```{r}
#Designate the validation set as 10% of the MovieLens data
set.seed(1, sample.kind="Rounding")
test_index <-createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <-movielens[-test_index,]
temp <-movielens[test_index,]
#Check if userId and movieId in validation set are also in edx set
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")
```

```{r}
#Merge the rows that were removed from the validation set back into the edx set
removed <-anti_join(temp, validation)
edx <-rbind(edx, removed)
rm(dl, ratings, movies, test_index, temp, movielens, removed)
```

```{r}
#divide Training and Test Sets:
set.seed(1, sample.kind = "Rounding")
test_index <-createDataPartition(y = edx$rating, times = 1, p = 0.1, list = F)
edx_train <-edx[-test_index,]
edx_temp <-edx[test_index,]
#Make sure userId and movieId are in the train and test sets
edx_test <-edx_temp %>%
  semi_join(edx_train, by = "movieId") %>%
  semi_join(edx_train, by = "userId")
removed <-anti_join(edx_temp, edx_test)
edx_train <-rbind(edx_train, removed)
rm(edx_temp, test_index, removed)
```
Analyzing the data
```{r}
edx %>% as_tibble()

#Confirm the dimensions and explore the features and classes of edx.
glimpse(edx)

#Determine the unique number of userIds, movieIds, and genres

edx %>% summarize(unique_users = length(unique(userId)),
                  unique_movies = length(unique(movieId)),
                  unique_genres = length(unique(genres)))

```

```{r}
#Ratings
length(unique(edx$rating))
```

```{r}
unique_ratings <-unique(edx$rating)
sort(unique_ratings)
```

```{r}
#View a Tibble of the Ratings Distribution
edx %>% group_by(rating) %>% summarize(ratings_sum = n()) %>%
  arrange(desc(ratings_sum))
```

```{r}
rp <-edx %>% filter(edx$rating >=3)
nrow(rp)/length(edx$rating)
```
TIMESTAMP
```{r}
#Transform the timestamp column of the edx dataset to a 'RatingYear' format
edx <- edx %>% mutate(timestamp = as.POSIXct(timestamp, origin = "1970-01-01", 
                                             tz = "EST"))
edx$timestamp <- format(edx$timestamp, "%Y")
names(edx)[names(edx) == "timestamp"] <- "RatingYear"
head(edx)
```

```{r}
validation <- validation %>% mutate(timestamp = as.POSIXct(timestamp, origin = "1970-01-01", 
                                                           tz = "EST"))
validation$timestamp <- format(validation$timestamp, "%Y")
names(validation)[names(validation) == "timestamp"] <- "RatingYear"
head(validation)

```

```{r}
edx_train <- edx_train %>% mutate(timestamp = as.POSIXct(timestamp, origin = "1970-01-01", 
                                                         tz = "EST"))
edx_train$timestamp <- format(edx_train$timestamp, "%Y")
names(edx_train)[names(edx_train) == "timestamp"] <- "RatingYear"
head(edx_train)
```

```{r}
edx_test <-edx_test %>% mutate(timestamp = as.POSIXct(timestamp, origin = "1970-01-01", 
                                                      tz = "EST"))
edx_test$timestamp <- format(edx_test$timestamp, "%Y")
names(edx_test)[names(edx_test) == "timestamp"] <- "RatingYear"
head(edx_test)
```

```{r}
range(edx$RatingYear)
```

```{r}
#Convert the 'RatingYear' column from character to numeric data type in order to plot a histogram
edx$RatingYear <-as.numeric(edx$RatingYear)
str(edx)
```

```{r}
edx %>% group_by(RatingYear, title) %>% 
  summarize(Ratings_Sum = n(), Average_Rating = mean(rating)) %>%
  mutate(Average_Rating = sprintf("%0.2f", Average_Rating)) %>%
  arrange(-Ratings_Sum) %>% print(n = 50)
```

```{r}
edx_genres <-edx %>% separate_rows(genres, sep = "\\|")
```

Sum of Movie Ratings per Genre

```{r}
edx_genres %>%
  group_by(genres) %>% summarize(Ratings_Sum = n(), Average_Rating = mean(rating)) %>%
    arrange(-Ratings_Sum)
```

```{r}
library(treemap)

# sum of  ratings by genre
genre_ratings <- aggregate(rating ~ genres, edx, sum)
# construct treemap
treemap(genre_ratings, index = "genres", vSize = "rating",
        type = "value", palette = "Set3",
        title = "Genre Ratings")
```

```{r}
#Arrange the Genres by Mean Rating
edx_genres %>%
  group_by(genres) %>% summarize(Ratings_Sum = n(), Average_Rating = mean(rating)) %>%
  arrange(-Average_Rating)
```

```{r}
#Coerce the 'genres' column from character data type to factor data type
edx$genres <-as.factor(edx$genres)
edx_genres$genres <-as.factor(edx_genres$genres)
class(edx_genres$genres)
```

```{r}
library(ggplot2)

# Aggregate of ratings per genre
genre_ratings <- edx %>%
  separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarize(total_ratings = sum(rating))

ggplot(genre_ratings, aes(x = reorder(genres, -total_ratings), y = total_ratings, fill = genres)) +
  geom_bar(stat = "identity") +
  ggtitle("Sum of Movie Ratings per Genre") +
  xlab("Genre") +
  ylab("Total Ratings") +
  theme(plot.title = element_text(hjust = 0.5))

```

Mean Rating per Genre

```{r}
library(ggplot2)

mean_ratings <- edx %>%
  separate_rows(genres, sep = "\\|") %>%
  group_by(genres) %>%
  summarize(mean_rating = mean(rating), .groups = 'drop')

ggplot(mean_ratings, aes(x = reorder(genres, mean_rating), y = mean_rating, fill = genres)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  ggtitle("Mean Rating per Genre") +
  xlab("Genre") +
  ylab("Mean Rating") +
  theme(plot.title = element_text(hjust = 0.5))

```


```{r}
yearreleaseda <-as.numeric(str_sub(edx$title, start = -5, end = -2))
edx <- edx %>% mutate(yearReleased = yearreleaseda)
head(edx)
```

```{r}
#Do the same for the validation set
yearreleasedb <-as.numeric(str_sub(validation$title, start = -5, end = -2))
validation <- validation %>% mutate(yearReleased = yearreleasedb)
head(validation)
```

```{r}
#This is also applied to edx_train & edx_test for later modeling purposes
yearreleasedc <-as.numeric(str_sub(edx_train$title, start = -5, end = -2))
edx_train <- edx_train %>% mutate(yearReleased = yearreleasedc)
head(edx_train)
```

```{r}
yearreleasedd <-as.numeric(str_sub(edx_test$title, start = -5, end = -2))
edx_test <- edx_test %>% mutate(yearReleased = yearreleasedd)
head(edx_test)
```

Use the newly defined “yearReleased” column to add a “MovieAge” column

```{r}
edx <-edx %>% mutate(MovieAge = 2020 - yearReleased)
validation <-validation %>% mutate(MovieAge = 2020 - yearReleased)
edx_train <-edx_train %>% mutate(MovieAge = 2020 - yearReleased)
edx_test <-edx_test %>% mutate(MovieAge = 2020 - yearReleased)
```

Movie Age

```{r}
summary(edx$MovieAge)
```
Modeling
The formula for RMSE can be defined as follows with y¯¯¯u,i the prediction of movie i by user u, and yu,the rating of movie i, by user u. N is then defined as the number of user/movie combinations and the sum of these different combinations.

```{r}
RMSE <- function(true_ratings, predicted_ratings){
  sqrt(mean((true_ratings - predicted_ratings)^2))
}

```

Begin Modeling: Benchmarking Model

```{r}
edx_train_mu <-mean(edx_train$rating)
NRMSE_M1 <- RMSE(edx_test$rating, edx_train_mu)
#Table the Results
results_table <-tibble(Model_Type = "NRMSE", RMSE = NRMSE_M1) %>% 
                  mutate(RMSE = sprintf("%0.4f", RMSE))
results_table
```
Median Table 

```{r}
edx_train_median <-median(edx_train$rating)
MM_M2 <-RMSE(edx_test$rating, edx_train_median)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model"),
                       RMSE = c(NRMSE_M1, MM_M2)) %>% 
                        mutate(RMSE = sprintf("%0.4f", RMSE))
results_table
```
Movie Effects Model

```{r}
bi <- edx_train %>% group_by(movieId) %>%
  summarize(b_i = mean(rating - edx_train_mu))
```

create the prediction 
```{r}
prediction_bi <-edx_train_mu + edx_test %>%
  left_join(bi, by = "movieId") %>% .$b_i
MEM_M3 <-RMSE(edx_test$rating, prediction_bi)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3)) %>% 
                        mutate(RMSE = sprintf("%0.4f", RMSE))
results_table

```

Adding User Effects to the Movie Effects Model:
```{r}
bu <-edx_train %>% left_join(bi, by = "movieId") %>% group_by(userId) %>%
  summarize(b_u = mean(rating - edx_train_mu - b_i))
```

Create the Prediction
Then check the prediction against the test set to determine the RMSE and table the results.
```{r}
prediction_bu <-edx_test %>% left_join(bi, by = "movieId") %>%
  left_join(bu, by = "userId") %>%
  mutate(predictions = edx_train_mu + b_i + b_u) %>% .$predictions
UEM_M4 <-RMSE(edx_test$rating, prediction_bu)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects", "Movie & User Effects"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3, UEM_M4)) %>% 
                        mutate(RMSE = sprintf("%0.4f", RMSE))
results_table
```

Adding Movie Age Effects:
(Movie, User & Movie Age Effects Model)
```{r}
ba <- edx_train %>%
  left_join(bi, by="movieId") %>% left_join(bu, by ="userId") %>%
  group_by(MovieAge) %>% summarize(b_a = mean(rating - b_i - b_u - edx_train_mu))
```

Create the Prediction
Check the prediction against the test set to determine the RMSE and table the results.

```{r}
predictions_ma <- edx_test %>% 
  left_join(bi, by = "movieId") %>% left_join(bu, by = "userId") %>%
  left_join(ba, by = "MovieAge") %>%  mutate(predictions = edx_train_mu + b_i + b_u + b_a) %>% 
  .$predictions
UMMAE_M5 <-RMSE(edx_test$rating, predictions_ma)
#Table the results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects", 
                                      "Movie & User Effects",
                                      "User, Movie & Movie Age Effects"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3, UEM_M4, UMMAE_M5)) %>% 
                        mutate(RMSE = sprintf("%0.4f", RMSE))
results_table
```
Movie & User Effects Model with Regularization:
```{r}
lambdasR <-seq(0, 10, 1)
RMSES <- sapply(lambdasR, function(l){
  edx_train_mu <- mean(edx_train$rating)
  
  b_i <- edx_train %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - edx_train_mu)/(n() + l))
  
  b_u <- edx_train %>%
    left_join(b_i, by='movieId') %>% 
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - edx_train_mu)/(n() +l))
  
  predicted_ratings <- edx_test %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    mutate(pred = edx_train_mu + b_i +  b_u) %>% .$pred
  
return(RMSE(predicted_ratings, edx_test$rating))
})
#Determine which lambda minimizes the RMSE
lambda <- lambdasR[which.min(RMSES)]
lambda
```

```{r}
library(ggplot2)

# Create a data frame with lambdasR and RMSES
data <- data.frame(lambdasR = lambdasR, RMSES = RMSES)

# Create the scatter plot
ggplot(data, aes(x = lambdasR, y = RMSES)) +
  geom_point(color = "purple", alpha = 0.3) +
  ggtitle("RMSE vs. Lambda") +
  xlab("Lambda") + ylab("RMSE")

```

Building the Movie & User Effects Model with Regularization
```{r}
b_i <- edx_train %>% 
  group_by(movieId) %>%
  summarize(b_i = sum(rating - edx_train_mu)/(n()+lambda))
b_u <-edx_train %>% 
  left_join(b_i, by="movieId") %>%
  group_by(userId) %>%
  summarize(b_u = sum(rating - b_i - edx_train_mu)/(n()+lambda))
reg_prediction <- edx_test %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  mutate(predictions = edx_train_mu + b_i + b_u) %>% .$predictions

UMEM_REG_M6 <-RMSE(edx_test$rating, reg_prediction)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects", 
                                      "Movie & User Effects",
                                      "Movie, User & Movie Age Effects",
                                      "Movie & User Effects w/Regularization"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3, UEM_M4, 
                                UMMAE_M5, UMEM_REG_M6)) %>% 
                        mutate(RMSE = sprintf("%0.6f", RMSE))
results_table
```
Movie, User & Movie Age Effects Model with Regularization:
```{r}
lambdasM <-seq(0, 10, 1)
RMSES2 <-sapply(lambdasM, function(l){
  edx_train_mu <-mean(edx_train$rating)
  
  b_i <-edx_train %>%
    group_by(movieId) %>%
    summarize(b_i = sum(rating - edx_train_mu)/(n() + l))
  
  b_u <-edx_train %>%
    left_join(b_i, by='movieId') %>% 
    group_by(userId) %>%
    summarize(b_u = sum(rating - b_i - edx_train_mu)/(n() +l))
  
  b_a <-edx_train %>% 
    left_join(b_i, by = "movieId") %>% left_join(b_u, by = "userId") %>%
    group_by(MovieAge) %>%
    summarize(b_a = sum(rating - b_i - b_u - edx_train_mu)/(n()+l))
  
  predicted_ratings <-edx_test %>%
    left_join(b_i, by = "movieId") %>%
    left_join(b_u, by = "userId") %>%
    left_join(b_a, by = "MovieAge") %>%
    mutate(predictions = edx_train_mu + b_i + b_u + b_a) %>% .$predictions
  
  return(RMSE(predicted_ratings, edx_test$rating))
})
lambda2 <- lambdasM[which.min(RMSES2)]
lambda2

```

```{r}
library(ggplot2)

# Create a data frame with lambdasM and RMSES2
df <- data.frame(lambda = lambdasM, RMSE = RMSES2)

# Create a scatterplot of RMSEs vs lambdasM
ggplot(df, aes(x = lambda, y = RMSE)) +
  geom_point() +
  geom_vline(xintercept = lambda2, linetype = "dashed") +
  ggtitle("RMSE vs lambda") +
  xlab("Lambda") + ylab("RMSE")

```
Building the User, Movie & Movie Age Effects Model with Regularization
```{r}
b_i <- edx_train %>% 
  group_by(movieId) %>%
  summarize(b_i = sum(rating - edx_train_mu)/(n()+lambda2))
b_u <-edx_train %>% 
  left_join(b_i, by = "movieId") %>%
  group_by(userId) %>%
  summarize(b_u = sum(rating - b_i - edx_train_mu)/(n()+lambda2))
b_a <-edx_train %>% 
  left_join(b_i, by = "movieId") %>% left_join(b_u, by = "userId") %>%
  group_by(MovieAge) %>%
  summarize(b_a = sum(rating - b_i - b_u - edx_train_mu)/(n()+lambda2))
reg_prediction2 <- edx_test %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_a, by = "MovieAge") %>%
  mutate(pred = edx_train_mu + b_i + b_u + b_a) %>%
  pull(pred)
UMMAE_REG_M7 <-RMSE(edx_test$rating, reg_prediction2)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects", 
                                      "Movie & User Effects",
                                      "User, Movie & Movie Age Effects",
                                      "Movie & User Effects w/Regularization",
                                      "User, Movie & Movie Age Effects w/Regularization"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3, UEM_M4, 
                                UMMAE_M5, UMEM_REG_M6, UMMAE_REG_M7)) %>% 
                        mutate(RMSE = sprintf("%0.5f", RMSE))
results_table
```
Using Validation:
Now we will move on to using the edx & validation sets to confirm our Final Model achieves an RMSE less than .8649.

The Benchmarking Model with Validation:
```{r}
edx_mu <-mean(edx$rating)
FRMSE_M1 <-RMSE(validation$rating, edx_mu)
#Table the Results
results_table <-tibble(Model_Type = ("NRMSE"),
                       Final_RMSE_Validation = (NRMSE_M1)) %>%
                      mutate(Final_RMSE_Validation = sprintf("%0.5f", Final_RMSE_Validation))
results_table
```
Median Model with validation:
```{r}
edx_med <-median(edx$rating)
FRMSE_M2 <-RMSE(validation$rating, edx_med)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2)) %>% 
  mutate(Final_RMSE_Validation = sprintf("%0.5f", Final_RMSE_Validation))
results_table
```

```{r}
bi <- edx %>% group_by(movieId) %>%
  summarize(b_i = mean(rating - edx_mu))
#Prediction
prediction_bi <-edx_mu + validation %>% 
  left_join(bi, by = "movieId") %>% .$b_i
FRMSE_M3 <-RMSE(validation$rating, prediction_bi)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, FRMSE_M3)) %>% 
  mutate(Final_RMSE_Validation = sprintf("%0.5f", Final_RMSE_Validation))
results_table
```
Movie & User Effects Model with Validation
```{r}
bu <-edx %>% left_join(bi, by = "movieId") %>% group_by(userId) %>%
  summarize(b_u = mean(rating - edx_mu - b_i))
#Prediction
prediction_bu <-validation %>% left_join(bi, by = "movieId") %>%
  left_join(bu, by = "userId") %>%
  mutate(predictions = edx_mu + b_i + b_u) %>% .$predictions
FRMSE_M4 <-RMSE(validation$rating, prediction_bu)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects",
                                      "Movie & User Effects"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, FRMSE_M3,
                                                 FRMSE_M4)) %>% 
  mutate(Final_RMSE_Validation = sprintf("%0.5f", Final_RMSE_Validation))
results_table
```

Movie, User & Movie Age Effects with Validation:
```{r}
ba <- edx %>%
  left_join(bi, by = "movieId") %>% left_join(bu, by = "userId") %>%
  group_by(MovieAge) %>% summarize(b_a = mean(rating - b_i - b_u - edx_mu))
#Prediction
predictions_ma <- validation %>% 
  left_join(bi, by = "movieId") %>% left_join(bu, by = "userId") %>%
  left_join(ba, by = "MovieAge") %>%  mutate(predictions = edx_mu + b_i + b_u + b_a) %>% 
  .$predictions
FRMSE_M5 <-RMSE(validation$rating, predictions_ma)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects",
                                      "Movie & User Effects",
                                      "Movie, User, & Movie Age Effects"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, FRMSE_M3,
                                                 FRMSE_M4, FRMSE_M5)) %>% 
  mutate(Final_RMSE_Validation = sprintf("%0.5f", Final_RMSE_Validation))
results_table
```
Movie & User Effects with Regularization (Validation):
```{r}
lambda
#Movie & User Effects Model with Regularization using the validation set

b_i <-edx %>% 
  group_by(movieId) %>%
  summarize(b_i = sum(rating - edx_mu)/(n()+lambda))
b_u <-edx %>% 
  left_join(b_i, by="movieId") %>%
  group_by(userId) %>%
  summarize(b_u = sum(rating - b_i - edx_mu)/(n()+lambda))
reg_prediction <-validation %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  mutate(predictions = edx_mu + b_i + b_u) %>% .$predictions

FRMSE_M6 <-RMSE(validation$rating, reg_prediction)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects",
                                      "Movie & User Effects",
                                      "Movie, User, & Movie Age Effects",
                                      "Movie & User Effects w/Regularization"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, FRMSE_M3,
                                                 FRMSE_M4, FRMSE_M5,
                                                 FRMSE_M6)) %>% 
                        mutate(Final_RMSE_Validation = sprintf("%0.5f", 
                                               Final_RMSE_Validation))
results_table
```

Final Model with Validation:
This Model features Movie, User, & Movie Age Effects with Regularization

```{r}
lambda2

b_i <- edx %>% 
  group_by(movieId) %>%
  summarize(b_i = sum(rating - edx_mu)/(n()+lambda2))
b_u <-edx %>% 
  left_join(b_i, by="movieId") %>%
  group_by(userId) %>%
  summarize(b_u = sum(rating - b_i - edx_mu)/(n()+lambda2))
b_a <-edx %>% 
  left_join(b_i, by="movieId") %>% left_join(b_u, by= "userId") %>%
  group_by(MovieAge) %>%
  summarize(b_a = sum(rating - b_i - b_u - edx_mu)/(n()+lambda2))
reg_prediction2 <-validation %>% 
  left_join(b_i, by = "movieId") %>%
  left_join(b_u, by = "userId") %>%
  left_join(b_a, by = "MovieAge") %>%
  mutate(predictions = edx_mu + b_i + b_u + b_a) %>% .$predictions

FRMSE_M7 <-RMSE(validation$rating, reg_prediction2)
#Table the Results
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects",
                                      "Movie & User Effects",
                                      "Movie, User, & Movie Age Effects",
                                      "Movie & User Effects w/Regularization",
                                      "Movie, User & Movie Age Effects w/Regularization"),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, FRMSE_M3,
                                                 FRMSE_M4, FRMSE_M5,
                                                 FRMSE_M6, FRMSE_M7)) %>% 
  mutate(Final_RMSE_Validation = sprintf("%0.5f", 
                                         Final_RMSE_Validation))
results_table
#Building the User, Movie & Movie Age Effects Model with Regularization

```

Table the training & test set results against those of the validation set
```{r}
results_table <-tibble(Model_Type = c("NRMSE", "Median_Model", "Movie Effects", 
                                      "Movie & User Effects",
                                      "Movie, User & Movie Age Effects",
                                      "Movie & User Effects w/Regularization",
                                      "User, Movie & Movie Age Effects w/Regularization"),
                       RMSE = c(NRMSE_M1, MM_M2, MEM_M3, UEM_M4, 
                                UMMAE_M5, UMEM_REG_M6, UMMAE_REG_M7),
                       Final_RMSE_Validation = c(FRMSE_M1, FRMSE_M2, 
                                                 FRMSE_M3, FRMSE_M4,
                                                 FRMSE_M5, FRMSE_M6,
                                                 FRMSE_M7)) %>%
  mutate(Final_RMSE_Validation = sprintf("%0.5f", 
                                         Final_RMSE_Validation)) %>%
  mutate(RMSE = sprintf("%0.5f", RMSE))
                       
                       
results_table
```

```{r}
#The kable function in knitr table of the final results
results_table %>% knitr::kable()
```

```{r}
# Create a data frame with the given data
model_data <- data.frame(
  Model_Type = c("NRMSE", "Median_Model", "Movie Effects", "Movie & User Effects",
                 "Movie, User & Movie Age Effects", "Movie & User Effects w/Regularization",
                 "User, Movie & Movie Age Effects w/Regularization"),
  RMSE = c(1.06005, 1.16676, 0.94296, 0.86468, 0.86433, 0.86414, 0.86384),
  Final_RMSE_Validation = c(1.06120, 1.16802, 0.94391, 0.86535, 0.86500, 0.86482, 0.86452)
)

# Create a bar plot of RMSE
barplot(
  model_data$RMSE,
  names.arg = model_data$Model_Type,
  xlab = "Model Type",
  ylab = "RMSE",
  main = "RMSE by Model Type",
  col = "blue"
)

# Create a bar plot of Final_RMSE_Validation
barplot(
  model_data$Final_RMSE_Validation,
  names.arg = model_data$Model_Type,
  xlab = "Model Type",
  ylab = "Final RMSE Validation",
  main = "Final RMSE Validation by Model Type",
  col = "red"
)



```


This Final Model achieves an RMSE of .86452
The lowest RMSE using the validation set is the Final Validation Model featuring Regularized User, Movie & Movie Age Effects.


